---
title: "R Cheat Sheets & Notes"
---

## Cheat Sheets

* [R Colors](http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf)

* [R Base Graphics Cheat Sheet](http://www.joyce-robbins.com/wp-content/uploads/2016/04/BaseGraphicsCheatsheet.pdf)

* [R Base Commands Cheat Sheet](https://www.rstudio.com/wp-content/uploads/2016/10/r-cheat-sheet-3.pdf)

* [R Markdown Cheat Sheet](https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf)

* [Keyboard Shortcuts](https://support.rstudio.com/hc/en-us/articles/200711853-Keyboard-Shortcuts)

## Notes

Add your own notes here...

```{r}
# Add R commands you want to remember here...

```

```{r}
# Or create other R chunks to put them in like this...

```

```{r, eval=FALSE}
# Using ```{r, eval=FALSE} turns off the chunk, but still shows it.
# Useful when you want to remember code, but not run it in this file.
```


```{r, eval=FALSE}

# Create the data:
set.seed(1140411)
sample1 <- rnorm(30, 69, 2.5)
sample2 <- rnorm(30, 69, 2.5)
theData <- data.frame(values = c(sample1,sample2), group = rep(c(1,2), each=30))
View(theData)
boxplot(values ~ group, data = theData)

 

# Run the permutation test:

myTest <-  t.test(values~group, data=theData, mu=0)
observedTestStat <- myTest$statistic
observedTestStat
 

N <- 2000      
permutedTestStats <-  rep(NA, N)
for  (i in 1:N ) {
   permutedData <- sample(x=theData$group)
   permutedTest <- t.test(values~permutedData,data=theData, mu=0)
   permutedTestStats[i]  <-  permutedTest$statistic
}
hist(permutedTestStats)
abline(v=observedTestStat)
sum(permutedTestStats >= observedTestStat)/N
sum(permutedTestStats <= observedTestStat)/N
sum(permutedTestStats <= observedTestStat)*2/N

```




set.seed(121)
sample1 <- rnorm(30, 185, 8)
sample2 <- sample1 - rnorm(30, 0, 3.5)
theData <- data.frame(values = c(sample1,sample2), group = rep(c(1,2), each=30), id = rep(c(1:30),times=2))
View(theData)
with(theData, hist(values[group==1] - values[group==2]))
```


mean(sample1)-mean(sample2)
mean(sample2)-mean(sample1)
```




# Create the data:

set.seed(121)
sample1 <- rnorm(30, 185, 8)
sample2 <- sample1 - rnorm(30, 0, 3.5)
theData <- data.frame(values = c(sample1,sample2), group = rep(c(1,2), each=30), id = rep(c(1:30),times=2))
View(theData)
with(theData, hist(values[group==1] - values[group==2]))

 

# Perform the permutation test:

myTest <-  t.test(values~group,data=theData, paried=TRUE,m=0)
observedTestStat <- myTest$statistic

 

N <- 2000      
permutedTestStats <-  rep(NA, N)
for  (i in 1:N ){
   permutedData <- sample(x=c(1,-1), size=30,replace=TRUE)
   
   permutedTest <- with(theData, t.test(permutedData * values[group==1]-values[group==2],mu=0)
   permutedTestStats[i]  <-  permutedTest$statistic
}

hist(permutedTestStats)
abline(v=observedTestStat)
sum(permutedTestStats >= observedTestStat)/N
sum(permutedTestStats <= observedTestStat)/N

```


## ANOVA AND REPLACEMENT EXAMPLE



  choc$Cocao <- as.numeric(gsub("%","",choc$Cocao))

Chocaov<-aov(Cocao~BroadOrig, data=choc)
  par(mfrow=c(1,2))
plot(Chocaov, which=1:2, pch=16)
```

## KRUSKAL WALLACE
**Make A Table Summary**

Friendly %>%
  group_by(condition) %>%
  summarise(min = min(correct), median = median(correct), mean = mean(correct), max = max(correct), sd = sd(correct), `Number of Students` = n()) %>%
pander(caption="Summary Of Results")
```

Normal QQ Plot


qqnorm(Friendly$correct)
qqline(Friendly$correct)
```

Strip Chart

stripchart(Friendly$correct~Friendly$condition, method="stack", xlab="Number Remembered Accurately of 40", main="Which Learning Method is the Best", ylab="Learning Method")

```

Wilcoxan Test


pander(
wilcox.test( correct ~ condition, data=F5, mu=0, alternative="two.sided" , confidence=0.95)
)
```

Filtering: for LM and Select


gp<-filter(OWD, hero=="Genji")
gp1<-filter(gp, stat_name=="Dragonblade Efficiency" | stat_name=="Eliminations" | stat_name=="Dragonblades")

gp2<-select(gp1, "stat_amount", "player", "stat_name", "start_time") %>% group_by(start_time, player) %>%
  pivot_wider(names_from = stat_name, values_from = stat_amount) %>% 
  rename(Dragonblade_Efficiency="Dragonblade Efficiency") %>% 
  mutate(Dragonblade_Efficiency=ifelse(is.na(Dragonblade_Efficiency), 0, Dragonblade_Efficiency)) %>% 
  mutate(Dragonblades=ifelse(is.na(Dragonblades), 0, Dragonblades))
```

Filtering: Linear Model Genji Example


genjilm <- lm(Eliminations ~ Dragonblade_Efficiency, data=gp2)
par(mfrow=c(1,3))
plot(genjilm, which=1)
qqPlot(genjilm$residuals, id=FALSE)
plot(genjilm$residuals, main="Residuals vs Order")
```


FB <- lm(Eliminations ~ Dragonblade_Efficiency, data=gp2)
pander(summary(FB))
```

plot(Eliminations ~ Dragonblade_Efficiency, data=gp2, pch=20, col="firebrick", cex=1.2, las=1,
     xlab="DragonBlade Efficcney", ylab="Eliminations",
     main="Eliminaitons Correlated with DragonBlade Efficiency")

genjilm <- lm(Eliminations ~ Dragonblade_Efficiency, data=gp2)
abline(genjilm, lwd=2, col=rgb(.689,.133,.133, .3))
abline(h=seq(0,120,20), v=seq(5,25,5), lty=2, col=rgb(.2,.2,.2,.2))

**LM HomeWise**

ProductDiversityLM<- lm(Total_Royalties ~ Unique_Product_Types, data=PDZ2)



## GG Plot Sideways Graph

Royalty2P<-Royaltiesbystore %>% ggplot(aes(x=reorder(Store, Total_Royalties_Earned), y=Total_Royalties_Earned), ylab="Quantity of Product Sold") + geom_col(fill="firebrick", color="black") + coord_flip()

Royalty2P + ggtitle("Stores Reveune") + ylab("Total Reveune") +xlab("Store")

**Linear Model Graph**

par(mfrow=c(1,1))
plot(Total_Royalties ~ Unique_Product_Types,  col= "red" , data=PDZ2, xlab="Unique Number of Product Types Sold By Store", ylab="Total Reveune Generated")
b<-ProductDiversityLM$coefficients
curve(b[1]+b[2]*x, add=TRUE, lwd=2, col='goldenrod')
curve((b[1] + b[3]) + (b[2] + b[4])*x, col="lightcoral", lwd=2, add=TRUE)


## Chi Squared Test of Independence


##Filtering the Data

GS2<-filter(GSS2012, pray %in% c(1,2,3,4,5,6))
       GS2<-filter(GSS2012, bible %in% c(1,2,3))
       
       GS2<-GS2 %>% 
         mutate( 
           bible=case_when(
             GS2$bible==1 ~ "Word Of God",
             GS2$bible==2 ~ "Inspired Word",
             GS2$bible==3 ~ "Book of Fables",
      
        )
         )
           
           GS2<-GS2 %>%  
             mutate(
           pray=case_when(
             GS2$pray==1 ~ "Several Times A day",
             GS2$pray==2 ~ "Once A Day",
             GS2$pray==3 ~ "Several times a week",
             GS2$pray==4 ~ "Once a Week",
             GS2$pray==5 ~ "Less than once a week",
             GS2$pray==6 ~ "Never",
         ))
```



prayert<-table(GS2$pray)

biblet<-table(GS2$bible)

praybiblet<-table(GS2$pray, GS2$bible)

  t2<-table(Pray=GS2$pray,Bible=GS2$bible)




pander(chi.t2$expected > 5)



pander(chi.t2 <- chisq.test(t2))
```


pander(mean(chi.t2$expected))
```

## Logisitic Regression

blow.glm <- glm(y>0 ~ D, data=blowAPB, family=binomial)
summary(blow.glm)
b<-coefficients(blow.glm)

**Plotting**

plot( y>0 ~ D, data=blowAPB, main="", ylab='Probability of Tree Dying', xlab="Tree Diameter in cm)", pch=16)
curve(exp(b[1]-b[2]*x)/(1-exp(b[1]-b[2]*x)), from=1, to=60, add=TRUE)
```

**Holsten Test (test of normality of logisitc regression**   

pander(hoslem.test(blow.glm$y, blow.glm$fitted, g=6))

##Kruskal Wallace

kruskal.test(choc2$Cocao, choc2$BroadOrig, data=choc2) %>% pander()

pander(favstats(Cocao ~ BroadOrig, data=choc2)[,-10])


## Wilcox Test:

pander(
wilcox.test( correct ~ condition, data=F5, mu=0, alternative="two.sided" , confidence=0.95)
)
