---
title: "The Sound of One Tree in the Forest"
output: 
  html_document: 
    theme: cerulean
    code_folding: hide
---

<hr />

#### Background

Trees Fall Over in storms, The point of this is to test which trees fall over and if there is one that is more likely to fall over than another.

In the data set.  The Trees Diameter was Recorded, if the tree died, and the tree type.
The Tree Types in this Data Set are Aspen and Paper Birch.

```{r, include=FALSE}
library(alr3)
library(mosaic)
library(DT)
library(dplyr)
library(pander)
```



```{r}
datatable(blowAPB)
```

<br />

#### Logistic Model

To model the probability of the a tree falling at diameters, we could apply the logistic regression model
$$
  P(Y_i = 1|x_i) = \frac{e^{\beta_0+\beta_1 x_i}}{1+e^{\beta_0 + \beta_1 x_i}} = \pi_i
$$
where for observation $i$: 

* $Y_i = 1$ denotes the tree falling, 
* $Y_i=0$ denotes the tree not falling, and 
* $x_i$ denotes the diameter of the tree. 

Note that if $\beta_1$ is zero in the above model, then $x_i$ (diameter) provides no insight about the probability of a tree dying. Thus, we could test the hypothesis that

$$
  H_0: \beta_1 = 0 \\
  H_a: \beta_1 \neq 0
$$

<br />


#### Exploring The Data

```{r}
blowAPB %>% group_by(SPP) %>% summarize(samplesize=n()) %>% pander()

```

As The Data Shows there are only two types of Tree Record in this Dataset.
Tree type A and Treee type PB.  And there are roughly the same number of each in the area studied.


#### Fitting the Model

To obtain estimates of the coefficients $\beta_0$ and $\beta_1$ for the `blowAPB` data, we apply the following R code.

```{r}
blow.glm <- glm(y>0 ~ D, data=blowAPB, family=binomial)
summary(blow.glm)
b<-coefficients(blow.glm)
```

Thus the estimated model for $\pi_i$ is given by
$$
  P(Y_i = 1|x_i) \approx \frac{e^{-2.280+0.1094 x_i}}{1+e^{-2.280+0.1094 x_i}} = \hat{\pi}_i
$$
where $b_0 = -2.280$ is the value of the `(Intercept)` which estimates $\beta_0$ and $b_1 = 0.1094$ is the value of `Diameter` which estimates $\beta_1$.

Importantly, the $p$-value for the test of `Diameter` shows a significant result $(p = 0.0320)$ giving sufficient evidence to conclude that $\beta_1 \neq 0$, which allows us to conclude that Diameter effects the probability of an A Tree falling.

<br />

#### Visualizing the Model

To visualize this simple logistic regression we could make the following plot.

```{r}
plot( y>0 ~ D, data=blowAPB, main="", ylab='Probability of Tree Dying', xlab="Tree Diameter in cm)", pch=16)
curve(exp(b[1]-b[2]*x)/(1-exp(b[1]-b[2]*x)), from=1, to=60, add=TRUE)
```

<br />

#### Diagnosing the Model

To demonstrate that the logistic regression is a good fit to these data we apply the Hosmer-Lemeshow goodness of fit test (since there are only a couple repeated $x$-values) from the `library(ResourceSelection)`. 

```{r, warning=FALSE, message=FALSE}
library(ResourceSelection)
```

```{r}
pander(hoslem.test(blow.glm$y, blow.glm$fitted, g=6))
# Note: doesn't give a p-value for g >= 7, default is g=10.
# Larger g is usually better than smaller g.
```

Recall that the null hypothesis is that the logistic regression is a good fit for the data, thus we reject the null, and claim that the logistic regression is not appropraite to predict this phenomina.

<br />

#### Conclusion

Since there is not a stable model, with tree diameter not being a prediction of falling over.  THere are other factors at place this model does not take into account, such as age of the tree, if it were infected, or already damaged, even its position in the forest. Thus unfortanetly this model cannot predict the lieklyood of a tree based on its diameter alone, as it does not have a group of like data.

Or it could be simiply saying that it doesnt matter how thick the trees are, they are always going fall over if they fall over. If there is something that will predict which trees will fall its not the diameter alone.

To be fair this is a mixed data set so perhaps just one tree is more resilant than another.
(How I checked that, the line was no different, althought the Hosmer-Lemeshow test came back as failing to reject the null in that case)


As far as I can tell the data is not a usable model and unless I missed something, we cannot run a prediction.

However should it have been so we could have done so using the Predict Function.

$$
  P(Y_i = 1|x_i) \approx \frac{e^{-2.280+0.1094\cdot (Tree Diameter)}}{1+e^{-2.280+0.1094 \cdot (Tree Diameter)}} = \hat{\pi}_i
$$
which, using R to do this calculation we get $\hat{\pi_i} \approx$

and used the code. 

predict(blow.glm, data.frame(D="Test Diameter Here"), type='response')

to predict the probaility of the tree falling over having that diameter.
As the model was not usable we unfortanetly were not able to use it to predict anything.  Not every test can be usable.
